Inefficient Concessions and Mediation

### 2017-06-22 presentation.tex
presentation at CesIfo Venice 06/15/17

\begin{document}
\title[Inefficient Concessions and Mediation\hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Inefficient Concessions and Mediation}
\author[Kristy Buzard and Ben Horne]{Kristy Buzard and Ben Horne\\ kbuzard@syr.edu}
\date{December 3, 2016}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\pause
\vskip.2in
Add time zero: let partners give concessions with both signaling and material value
\pause
\begin{itemize}[<+->]
	\item Let partners destroy some/all of the material value
	\item Let material value provide help/harm to the giver
\end{itemize}

@@

\section{Model}
\subsection{}
\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Countries simultaneously give costly concessions: $g_1, g_2 \in \mathbb{R}_+$
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}

@@

\begin{frame}
\begin{itemize}[<+->]
	\item Payoffs: sum the discounted stage game payoffs plus any concessions 
		\begin{itemize}
			\item e.g. player's $i$'s payoff if both parties play ``Trust'' in every period: $\sum_{t=1}^\infty \de_i^{t-1} T = \frac{T}{1-\de_i}$
		\end{itemize}
	\item Parameters are common knowledge with the exception of $\delta_i$, which is country i's private information
	\item Social welfare measured as sum of high types' expected utilities
\end{itemize}
\end{frame}

@@

\begin{frame}{Benchmark Model}
\pause
Assume two types: $\de_h$ and $\de_l$
\begin{itemize}[<+->]
	\item $\de_h > \de^* > \de_l$ where $\de^*$ is the cutoff for sustaining (Trust,Trust) eqm
	\item $p$: probability of high type
	\item Low type's cost to give concession $g$: $c_l(g)>c_h(g) =g$
\end{itemize}

\pause
\vskip.2in
Some equilibria of interest
\pause
\begin{itemize}[<+->]
	\item Pool on `Fight'
	\item Separating without concessions
	\item Separating through concessions
\end{itemize}
\end{frame}

@@

\pause	
\[
  U_h=(p-1)g + \textstyle \frac{\de_h}{(1-\de_h)} \left[pT+(1-p)(W-D)\right]
\]
\pause
\vskip-.1in

	\begin{itemize}[<+->]
		\item The smallest concession is $c_l^{-1}\left(p(T+D)\right)$
		\item If $p$ is low, high types are better off in the `fight' pooling equilibrium
	\end{itemize}
\end{frame}

\section{Inefficient Concessions}
\subsection{Add Money Burning}
\begin{frame}{Burning money is unattractive}
\pause
Allow giver of concession to also choose $0 \leq e \leq 1$, where benefit of concession is $eg$

\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 3}
  When concessions confer no material help or harm on the giver, it is optimal to give efficient gifts ($e=1$) in a separating eqm.
\end{beamerboxesrounded}
\pause
\begin{itemize}[<+->]
	\item The benefit of the gift appears on both sides of the incentive constraint for both individuals, so cancels out
	\item The benefit appears in the high type's expected utility
\end{itemize}
\end{frame}

\begin{frame}[label=hurt]
\frametitle{Concessions can hurt the giver}
\pause
Now assume a low type will use concessions against you, i.e. to reduce your payoffs when partner fights \hyperlink{Supplemental}{\beamerbutton{Modified Payoffs}}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 4}
  There are parameters under which the optimal equilibrium features inefficient concessions.
\end{beamerboxesrounded}
\pause
\[
  U_h=peg-g + \textstyle \frac{\de_h}{(1-\de_h)} \left[pT(1+eg)+(1-p)(W-D(1+eg))\right]
\]
\pause
\vskip-.1in
\begin{itemize}[<+->]
	\item If $p$ is low, concessions likely to be used against you, so remove material value
	\item Conjecture: Pooling on `Fight' optimal over larger set of parameters
\end{itemize}
\end{frame}

\section{Mediation}
\subsection{}
\begin{frame}{Mediator as mechanism designer}
\pause
`Manipulative' mediator: parties report their types, must deliver the stipulated concessions
\pause
\begin{itemize}[<+->]
  \item Mechanism: if two high types, give concession and play `Trust'; Otherwise, no concession and `Fight'
	\item Concession is necessary to get truthful revelation, but only to high type
	\item Cost of concession must be greater than benefit for low type
\end{itemize}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 5}
  A mediator can eliminate inefficient concessions.
\end{beamerboxesrounded}

>>>>

\begin{document}
\title[Inefficient Concessions and Mediation\hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Inefficient Concessions and Mediation}
\author[Kristy Buzard and Ben Horne]{Kristy Buzard and Ben Horne\\ kbuzard@syr.edu}
\date{June 15, 2017}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\pause
\vskip.2in
Add time zero: let partners give concessions with both signaling and material value
\pause
\begin{itemize}[<+->]
	\item Let material value provide help/harm to the giver
	\item Let partners destroy some/all of the material value
\end{itemize}

@@

\section{Model}
\subsection{}
\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Countries simultaneously give costly concessions: $g_i \in \mathbb{R}_+$
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}

@@

\begin{frame}
\begin{itemize}[<+->]
	\item Payoffs: sum the discounted stage game payoffs plus any concessions given or received
		\begin{itemize}
			\item e.g. player's $i$'s payoff if both parties give no concession and play ``Fight'' in every period: $$\sum_{t=1}^\infty \de_i^{t-1} (W-D) = \frac{W-D}{1-\de_i}$$
		\end{itemize}
	\item Parameters are common knowledge with the exception of $\delta_i$, which is country i's private information
	\item Social welfare measured as sum of high types' expected utilities
\end{itemize}
\end{frame}

@@

\begin{frame}{Benchmark Model}
\pause
Assume two types: $\de_h$ and $\de_l$
\begin{itemize}[<+->]
	\item $\de_h > \de^* > \de_l$ where $\de^*=\frac{W}{T+D}$ is the cutoff for sustaining (Trust,Trust) eqm
	\item $p$: probability of high type
	\item Cost of giving concessions $g$: $g = c_l(g)\geq c_h(g) = g$
\end{itemize}

\pause
\vskip.2in
Some equilibria of interest
\pause
\begin{itemize}[<+->]
	\item Separating without concessions
	\item Separating through concessions
	\item Pool on `Fight'
\end{itemize}
\end{frame}

@@

\pause	
\[
  U_h=pg - c_h(g) + \textstyle \frac{\de_h}{(1-\de_h)} \left[pT+(1-p)(W-D)\right]
\]
\pause
\vskip-.1in

	\begin{itemize}[<+->]
		\item The smallest concession is $p(T+D)$
		\item If $p$ is low, high types are better off in the `fight' pooling equilibrium
	\end{itemize}
\end{frame}

\section{Inefficient Concessions}
\subsection{Add Money Burning}
\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Concessions
		\begin{enumerate}
			\item[0a.] Countries simultaneously give costly concessions and decide on the efficiency of the concessions: $(g_i, e_i) \in \mathbb{R}_+ \times [0,1]$ 
			\item[0b.] \textcolor{red!75!green!50!blue!25!gray}{Countries decide what proportion of a received concession to invest in civil society (vs. military capabilities): $\alpha_i \in [0,1]$}
		\end{enumerate}	
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}
\end{frame}

\begin{frame}{Burning money is unattractive}
\pause
Now benefit of concession is $eg$

\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 3}
  When concessions confer no material help or harm on the giver, it is optimal to give efficient gifts ($e=1$) in a separating eqm.
\end{beamerboxesrounded}
\pause
\begin{itemize}[<+->]
	\item The benefit of the gift appears on both sides of the incentive constraint for both individuals, so cancels out
	\item Costs of giving a concession don't change
	\item The benefit appears in the high type's expected utility
\end{itemize}
\end{frame}

\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Concessions
		\begin{enumerate}
			\item[0a.] Countries simultaneously give costly concessions and decide on the efficiency of the concessions: $(g_i, e_i) \in \mathbb{R}_+ \times [0,1]$ 
			\item[0b.] \textcolor{blue}{Countries decide what proportion of a received concession to invest in civil society (vs. military capabilities): $\alpha_i \in [0,1]$}
		\end{enumerate}	
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}
\end{frame}

\begin{frame}[label=hurt]
\frametitle{Concessions can hurt the giver}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 4}
  When the low type country can use concessions to reduce the payoffs of its negotiating partner during a `fight' stage (i.e. $\alpha_L < 1$), there are parameters under which the separating-through-concessions equilibrium features inefficient concessions.
\end{beamerboxesrounded}
\hyperlink{Supplemental}{\beamerbutton{Modified Payoffs}}
\pause
\[
  U_h=peg - c_h(g) + \textstyle \frac{\de_h}{1-\de_h} \left[pT(1+eg)+(1-p)(W-D(1+eg))\right]
\]
\pause
\vskip-.1in
\begin{itemize}[<+->]
	\item If $p$ is low, concessions likely to be used against you, so remove material value
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Peace not possible in some scenarios}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 5}
  When $\alpha_L < 1$ is allowed, peace becomes unachievable for some parameters under which it is achievable when concessions do not have material value.
\end{beamerboxesrounded}

\pause
\vskip.1in
\begin{itemize}[<+->]
	\item If peace is achievable, concessions may be either efficient or inefficient
	\item High-type utility may increase or decrease from no-material-value case
\end{itemize}
\end{frame}

\section{Mediation}
\subsection{}
\begin{frame}{Mediator as mechanism designer}
\pause
`Manipulative' mediator: parties report their types, must deliver the stipulated concessions
\pause
\begin{itemize}[<+->]
  \item Mechanism: if two high types, give concession and play `Trust'; Otherwise, no concession and `Fight'
	\item Concession is necessary to get truthful revelation, but only to high type
	\item Need cost of concession for low type to be not to large relative to cost for high type
\end{itemize}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 6}
  A mediator restores peace where $\alpha_i < 1$ destroys it, eliminates inefficient concessions, and reduces the threshold $\de_h$ for high types.
\end{beamerboxesrounded}
